---
title: "DSC520_Week4_Guruprasad_Velikadu_Krishnamoorthy_Part_4.1"
author: "Guruprasad Velikadu Krishnamoorthy"
date: "2023-01-08"
output: 
  pdf_document: 
    highlight: espresso
---
# Assignment Week 4 Part 1
### Loading the required Packages
```{r}
library(ggplot2)
library(psych)
library(qqplotr)
library(pastecs)
```
### Set the working directory to the root of your DSC 520 directory and initial settings
```{r}
knitr::opts_knit$set(root.dir ="C:/Users/Gurup/GURU/Learning/Masters/Term_2/DSC520_T302_Statistics_for_Data_Science/Week_4/")
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```
### 4.Test Scores
## A professor has recently taught two sections of the same course with only one difference between the sections. In one section, he used only examples taken from sports applications, and in the other section, he used examples taken from a variety of application areas. The sports themed section was advertised as such; so students knew which type of section they were enrolling in. The professor has asked you to compare student performance in the two sections using course grades and total points earned in the course. You will need to import the Scores.csv dataset that has been provided for you.
## 4.1. What are the observational units in this study?
#### Solution : 
#####       The Observational unit is defined as the unit on which the statistics are collected for the dataset. Using the str function we can see there are 38 observations of 3 variables- Count, Score and Section which are the Observational units.
```{r}
scores_df <- read.csv(file="data/scores.csv",header=TRUE,stringsAsFactors=FALSE)
str(scores_df)
dim(scores_df)
```
### 4.2 Identify the variables mentioned in the narrative paragraph and determine which are categorical and quantitative?
#### Solution:
#####       The Categorical variable is a variable based on which the data can be split into various groups. Based on the scores.csv data, The Categorical Variable will be the "Section" that can categorize the data into "Sports" and "Regular" sections. The Quantitative variable provides the measurement and in our dataset, the Quantitative variable are "Score" that gives the measurement of the score by students and also the "Count" which is the number of students that received the score for each category. The results of SUmmary and Table function shows that there are 2 Unique values in the Section which can be the Categorical variable and the fields Score and Count are Numeric that are Quantitative.
```{r}
summary(scores_df)
table(scores_df$Section,scores_df$Score)
table(scores_df$Section,scores_df$Count)
table(scores_df$Score,scores_df$Count)
```
### 4.3 Create one variable to hold a subset of your data set that contains only the Regular Section and one variable for the Sports Section.
```{r}
regularData <- subset(scores_df,scores_df$Section=="Regular")
head(regularData)
sportsData <- subset(scores_df,scores_df$Section=="Sports")
head(sportsData)
```
### 4.4 Use the Plot function to plot each Sections scores and the number of students achieving that score. Use additional Plot Arguments to label the graph and give each axis an appropriate label. 
#### Solution:
#####       This section is answered by plotting Scatter plots using both Plot function and ggplot. 
```{r}
# Scatter Plot for Regular section scores
plot(regularData$Score,regularData$Count,col="Red",pch=16,main="Scores for the Regular Section",xlab="Scores",ylab="Number of Students")
# Scatter Plot for Sports section scores
plot(sportsData$Score,sportsData$Count,col="Blue",pch=17,main="Scores for the Sports Section",xlab="Scores",ylab="Number of Students")
# creating groups to be used in plots
group <-   as.factor(scores_df$Section)
# Scatter plot to show the two sections in one plot
plot(scores_df$Score,scores_df$Count,col=group,pch=16,main="Scores achieved by students in each Sections",xlab="Scores",ylab="Number of students")
legend("topleft",legend = c("Regular", "Sports"),col = 1:2,pch = 16)
```

```{r}
# Using ggplot 
ggplot(data=scores_df,aes(x=Score,y=Count))+geom_point(aes(col=Section),size=3)+labs(title ="Scores achieved by students in each Sections",x="Scores",y="Number of students")+theme_bw()
```

#### 4.4.a) Comparing and contrasting the point distributions between the two section, looking at both tendency and consistency: Can you say that one section tended to score more points than the other? Justify and explain your answer.
##### Solution: From the plot above ,it appears the students scored more in the Regular section than the Sports section. The range of scores in Sports section(195) is wider than the range in Regular(115) section. The Minimum score in Sports section is 200 while the minimum in Regular is 265.
```{r}
# creating a new dataframe to compare the consistency between scores
new_scores_df = data.frame(regularData$Score,sportsData$Score)
alpha(new_scores_df[,c("regularData.Score","sportsData.Score")])
```
##### From the output of the Consistency alpha function, the value of raw_data is 0.91 which represents strong consistency between the scores in the 2 sections. The tendency results based on mean shows that students scored better in Regular section that Sports.

#### 4.4.b)Did every student in one section score more points than every student in the other section? If not, explain what a statistical tendency means in this context.
##### Solution : To answer this question, we have to compare the tendency(Mean, Median) of Scores from each section. QQPlots and Normal Distribution curves were plotted to justify the solution with the plots.The Mean and Median of Regular scores is greater than Sports scores(Refer output of stat.desc function). Though it is not possible to compare one score against another score, overall in terms of comparing them against the Mean and median of each section and from the qq-plots, we can say Regular section fared better than Sports overall.

```{r}
# qqplot for Scores in Regular Section
ggplot(data = regularData["Score"], aes(sample = Score)) +stat_qq_point(size = 2, color = "red", distribution = "norm") +stat_qq_line(color = "blue") + labs(x = "Score", y="Sample",title = "QQ Plot for Scores in Regular section") +theme_bw()
# qqplot for Scores in Sports Section
ggplot(data = sportsData["Score"], aes(sample = Score)) +stat_qq_point(size = 2, color = "green", distribution = "norm") +stat_qq_line(color = "red") + labs(x = "Score",y = "Sample", title = "QQ Plot for Scores in Regular section") +theme_bw()
# qqplot for Scores for each Section
ggplot(data = scores_df, aes(sample = Score,color = Section)) +
    stat_qq_point(size = 2,  distribution = "norm") +stat_qq_line()+facet_wrap(~scores_df$Section)+ labs(x = "Score",                                                          y = "Sample", title = "Scores achieved by students")+theme_bw()+theme(legend.position="bottom")
# Histogram and Normal Curve for Regular Data
hist.regular <- ggplot(regularData,aes(Score))+geom_histogram(bins=10,aes(y=after_stat(density)),
                                                              color="black",fill="pink")+labs(x="Score",y="Density",title="Distribution for Regular section Scores")
hist.regular+stat_function(fun=dnorm,args=list(mean=mean(regularData$Score,na.rm=TRUE),sd=sd(regularData$Score,na.rm=TRUE)),color="blue",linewidth=1.5)
# Histogram and Normal Curve for Sports Data
hist.sports <- ggplot(sportsData,aes(Score))+geom_histogram(bins=10,aes(y=after_stat(density)),
                                                            color="black",fill="grey")+labs(x="Score",y="Density",title="Distribution for Sports section Scores")

hist.sports+stat_function(fun=dnorm,args=list(mean=mean(sportsData$Score,na.rm=TRUE),sd=sd(sportsData$Score,na.rm=TRUE)),color="Red",linewidth=1.5)
# Statistics for Scores from each section
stat.desc(regularData$Score,norm=TRUE)
stat.desc(sportsData$Score,norm=TRUE)
```

#### 4.4.c) What could be one additional variable that was not mentioned in the narrative that could be influencing the point distributions between the two sections?
##### Solution: The Student's grade in the dataset would have influences the point distributions in the plots. Instead of plotting the performance based on the scores which is hard to compare between each section, the Student's grade would have been an easy measure to compare the performane of students.For instance, with the additional data of Grades, we can tell how many students scored A in Regular vs Sports and that can tell how students fared in each section. With just the scores, there is nothing to compare with. 

# Session info


```{r}
sessionInfo()
```
